Visualizing demographic evolution using geographically inconsistent census data

### 2018-08-16 paper/study.tex
almost done expert

\section{Expert feedback}
%This could be a big step toward better understanding of evolving urban morphology.
To assess the proposed method, we contacted academic and industry experts in
sociology and urban sciences. The supplementary material contains the
communication with five experts, verbatim, identified from A to E.

@@

While the interface was "easy to navigate" (B), it was also considered
"overwhelming" (A), "intimidating" (E), and "tricky to interpret" (C), possible
side-effects of our effort to increase  representational accuracy, where we
avoided using simplified representation or labels. Identifying clusters by their
most relevant variables was welcome, but the overlap of information from
different clusters in the boxplot was "a bit confusing" (C) when color is not
present. Further, most clusters can be sufficiently characterized using only the
most relevant aspect, but this is not generally true.

While the map of trajectories was mentioned as a "good summary map", how it
related to the clustering method was unclear (C). While we provide different
options on how the colors are used, both are sub optimal since reliably
representing several distinct entities using colors is humanly unfeasible.
Indeed, the number of distinguishable colors was a significant constraint, we
found indications that more clusters should be used in some cases, even if eight
clusters is more than what is traditionally considered in these analyses.
Conversely, increasing the number of clusters would also complicate the
interpretation of the results.

The experts also mentioned the poor responsiveness of the method when changes in
the clustering parameters required server-side processing (B,D). Indeed, the
current implementation can take a few minutes to cluster regions with high
number of CTs, like Los Angeles or Brooklyn. Server-processing reduced the
amount of data transfered to client, but might increase response time under
load. We implemented a cache policy that greatly reduces the response time, but
fully pre-processing the results is not practical due to size of the parameter
space.

Most of the experts demonstrated interest in using our method in their research
(A, B, D, E), aiming to use the census data as a backdrop for other datasets,
providing demographic context. They also mentioned the need to export subsets of
data, plots, and maps to be used in reports and publications (C, D, E).

More importantly, while these experts were aware that our method does not
perform geographic normalization, none of them mention it. We did not
specifically ask if this difference led to unexpected results, but rather
inquired if they found interesting insights, leading them to consider the
interface, its options, and results as a whole. Most experts found phenomena
corroborated by the specialized literature, indicating that our methodology
produces equivalent results, with a fraction of the effort. We interpret the
fact that most of them were interested in the next steps as an affirmation of
the accuracy of the method.

>>>>

\section{Expert feedback}
%This could be a big step toward better understanding of evolving urban morphology.
To assess the proposed method, we contacted academic and industry experts in
sociology and urban sciences. We focused our inquires on the utility of the
prototype, and if they found anything interesting on the data using it. The
supplementary material contains the communication with five experts, verbatim,
identified from A to E.

@@

The interface was "easy to navigate" (B), but it was also considered
"overwhelming" (A), "intimidating" (E), and "tricky to interpret" (C), possible
side-effects of our effort to increase  representational accuracy, where we
avoided using simplified representation or labels. Identifying clusters by their
most relevant variables was welcome, but the overlap of information from
different clusters in the boxplot was "a bit confusing" (C) when color was not
present. Further, most clusters can be sufficiently characterized using only the
most relevant aspect, but this is not generally true.

While the map of trajectories was mentioned as a "good summary map", how it
related to the clustering method was unclear (C). The methos includes different
options on how the colors are used, but both are sub optimal since reliably
representing several distinct entities using colors is humanly unfeasible.
Indeed, the number of distinguishable colors was a significant constraint, we
found indications that more clusters should be used in some cases, even if eight
clusters is more than what is traditionally considered in these analyses.
Conversely, increasing the number of clusters would also complicate the
interpretation of the results.

The experts also mentioned the poor responsiveness of the method when changes in
the clustering parameters required server-side processing (B,D). Indeed, the
current implementation can take a few minutes to cluster regions with high
number of CTs, like Los Angeles or Brooklyn. Server-processing reduced the
amount of data transfered to client, but it might increase the response time
under load. We implemented a cache policy that greatly improved the performance,
but fully pre-processing the results is not practical due to size of the
parameter space.

Most of the experts demonstrated interest in using our method in their research
(A, B, D, E), aiming to use the census data as a backdrop for other datasets,
providing demographic context. They also mentioned the need to export subsets of
data, plots, and maps to be used in reports and publications (C, D, E). More
importantly, while these experts were aware that our method does not perform
geographic harmonization, none of them mention it. We did not specifically ask
if this difference led to unexpected results, but rather if they found
interesting insights. Most experts found phenomena corroborated by the
specialized literature, indicating that our methodology produces equivalent
results, with a fraction of the effort. We interpret the fact that most of them
were interested in the next steps as confirmation of the accuracy of the method.
